Language and vision in conceptual processing: Multilevel analysis and statistical power

### 2022-10-24 manuscript/manuscript.Rmd
24102022

  This manuscript is a draft and includes some appendices. Correspondence can be addressed to Pablo Bernabeu on pcbernabeu@gmail.com. All materials are available at [http://doi.org/10.17605/OSF.IO/UERYQ](http://doi.org/10.17605/OSF.IO/UERYQ).

abstract: |
   Research over the past two decades has suggested that conceptual processing depends on both language-based and vision-based information. We tested this interplay at three levels of the data: individuals, words and tasks. To this aim, we drew on three existing data sets that were composed of large samples, and implemented the paradigms of semantic priming, semantic decision and lexical decision. After extending these data sets with language-based and vision-based measures, we performed the analysis using mixed-effects models that included a comprehensive array of fixed effects---including covariates---and random effects. Overall, language-based information was found to be more important than vision-based information. Furthermore, in the semantic priming study---whose task required distinguishing between words and nonwords---, both language-based and vision-based information were more influential when words were presented faster. In addition, higher-vocabulary participants presented a greater effect of language-based similarity than lower-vocabulary participants. In contrast, this pattern did not hold for vision-based information, which was less relevant to the task. The 'relevance advantage' in higher-vocabulary participants also appeared in the semantic decision and the lexical decision studies. Furthermore, we compare the effects of two visual information measures, and discuss the role of measurement instruments in this research topic. Lastly, we estimated the sample size required to reliably examine each effect of interest. We found that 300 participants suffice to examine language-based information contained in words, whereas more than 1,000 participants are necessary to examine vision-based information as well as the interactions between both the former variables and vocabulary size, gender and presentation speed.

@@

output:
  papaja::apa6_pdf:
    extra_dependencies: ["flafter"]   # Place figures and tables below where they are first mentioned in the text.

# Latex preamble specifying layout
header-includes: 
# Background on the first four lines: https://stackoverflow.com/a/70658799/7050882
  - \renewcommand\author[1]{}
  - \renewcommand\affiliation[1]{}
  - \authorsnames[1, 2, 2]{Pablo Bernabeu, Dermot Lynott, Louise Connell}
  - \authorsaffiliations{{Department of Psychology, Lancaster University, UK}, {Department of Psychology, Maynooth University, Ireland}}
  - \renewcommand{\topfraction}{.75}        # Reduce page margins
  - \renewcommand{\bottomfraction}{.4}      # Reduce page margins
  - \renewcommand{\textfraction}{.15}       # Reduce page margins
  - \renewcommand{\floatpagefraction}{.6}   # Reduce page margins
  - \setlength{\@fptop}{0pt}                # Vertically-align tables & figures to the top
  - \setcounter{topnumber}{3}               # Reduce page margins
  - \setcounter{bottomnumber}{2}            # Reduce page margins
  - \setcounter{totalnumber}{4}             # Reduce page margins
  - \usepackage{amsmath}                    # Maths formatting
  - \usepackage{upgreek}                    # Disable the italicisation of Greek letters
  - \usepackage{booktabs}                   # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{longtable}                  # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{array}                      # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{multirow}                   # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{wrapfig}                    # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{float}                      # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{colortbl}                   # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{pdflscape}                  # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{tabu}                       # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{threeparttable}             # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{threeparttablex}            # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage[normalem]{ulem}             # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{makecell}                   # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{xcolor}                     # Dependency for R package 'kableExtra' (table formatting)
  - \renewcommand\appendix{}                # Remove redundant 'Appendix' label
---

@@

library(papaja)       # APA-formatted manuscript
library(knitr)        # Document rendering
options(kableExtra.latex.load_packages = FALSE)  # instead, LaTeX packages loaded in header/preamble above
library(kableExtra)   # Tables
library(dplyr)        # Data wrangling
library(formattable)  # Format numbers
library(kableExtra)   # Table formatting (e.g., `pack_rows()`)
library(stringr)      # Text processing
library(ggplot2)      # Plots
library(GGally)       # Correlation plots
library(sjPlot)       # Model plots
library(RColorBrewer) # Colours in plots
library(patchwork)    # Combination of plots
library(Cairo)        # Allows use of special characters such as dashes in plots
library(magick)       # Image rendering
library(tikzDevice)   # Image rendering

# Move working directory to the project root
knitr::opts_knit$set(root.dir = '../')

@@

# Subset of the semantic priming study that included visual similarity.
# Code for data set below in 'semanticpriming/data' folder
semanticpriming_with_visualsimilarity = 
  read.csv('semanticpriming/data/subset_with_visualsimilarity/semanticpriming_with_visualsimilarity.csv')

@@

# Code for models below in 'semanticpriming/bayesian_analysis' folder
semanticpriming_summary_informativepriors_exgaussian =   # Primary model object
  readRDS('semanticpriming/bayesian_analysis/results/semanticpriming_summary_informativepriors_exgaussian.rds')
# semanticpriming_summary_weaklyinformativepriors_exgaussian =   # Primary model object
#   readRDS('semanticpriming/bayesian_analysis/results/semanticpriming_summary_weaklyinformativepriors_exgaussian.rds')
semanticpriming_summary_diffusepriors_exgaussian =   # Primary model object
  readRDS('semanticpriming/bayesian_analysis/results/semanticpriming_summary_diffusepriors_exgaussian.rds')

@@

# Code for models below in 'semanticdecision/bayesian_analysis' folder
semanticdecision_summary_informativepriors_exgaussian =   # Primary model object
  readRDS('semanticdecision/bayesian_analysis/results/semanticdecision_summary_informativepriors_exgaussian.rds')
semanticdecision_summary_weaklyinformativepriors_exgaussian =   # Primary model object
  readRDS('semanticdecision/bayesian_analysis/results/semanticdecision_summary_weaklyinformativepriors_exgaussian.rds')
semanticdecision_summary_diffusepriors_exgaussian =   # Primary model object
  readRDS('semanticdecision/bayesian_analysis/results/semanticdecision_summary_diffusepriors_exgaussian.rds')

@@

# Code for models below in 'lexicaldecision/bayesian_analysis' folder
lexicaldecision_summary_informativepriors_exgaussian =   # Primary model object
  readRDS('lexicaldecision/bayesian_analysis/results/lexicaldecision_summary_informativepriors_exgaussian.rds')
lexicaldecision_summary_weaklyinformativepriors_exgaussian =   # Primary model object
  readRDS('lexicaldecision/bayesian_analysis/results/lexicaldecision_summary_weaklyinformativepriors_exgaussian.rds')
lexicaldecision_summary_diffusepriors_exgaussian =   # Primary model object
  readRDS('lexicaldecision/bayesian_analysis/results/lexicaldecision_summary_diffusepriors_exgaussian.rds')

@@

Over the past decades, research in psycholinguistics has suggested that conceptual processing depends on both language and embodiment systems [@barsalouLanguageSimulationConceptual2008; @louwerseSymbolInterdependencySymbolic2011; @connell2013a]. A comprehensive approach to both these systems requires a direct comparison of them. Studies tackling such a comparison have found that the systems are selectively engaged, following contextual demands [@connell2013a; @louwerseTasteWordsLinguistic2011; @ostarekTaskdependentCausalRole2017; @petilli2021a]. For instance, the role of perceptual simulation appears to be augmented in slower responses [@louwerseTasteWordsLinguistic2011], when words are presented more slowly [@lam2015a], and in tasks prompting deeper semantic processing [@connell2013a; @louwerseTasteWordsLinguistic2011; @ostarekTaskdependentCausalRole2017; @petilli2021a].

In spite of the amount of evidence favouring the interplay between language and embodiment, there are at least 2 reasons to continue testing the interplay theory. Firstly, the coexistence of several systems in a scientific theory must be thoroughly justified, due to the value of simplicity, as epitomised by Occam's razor [@galleseBrainConceptsRole2005; @tillmanHowSharpOccam2015]. This point is particularly pressing because the language system has consistently produced larger effect sizes than the embodiment system [@lam2015a; @pecherDoesPizzaPrime1998; @petilli2021a; @banksLinguisticDistributionalKnowledge2021; @kiela2014a; @louwerse2015a]. Consequently, in this research question, we examine whether the interplay between linguistic and visual processing holds in the expected directions and at multiple levels of the experimental structure.

Secondly, mixed evidence has appeared regarding some aspects of the interplay. For instance, whereas some studies have suggested that the language system is activated before the embodiment system [@lam2015a; @louwerseTasteWordsLinguistic2011], a recent study found the opposite pattern in a lexical decision task [@petilli2021a]. Similarly, some evidence has suggested that high-vocabulary participants are more sensitive to linguistic features [@yap2017a], whereas other evidence has suggested the opposite [@yapIndividualDifferencesJoint2009; @yap2012a]. Another case of mixed evidence regards gender: whereas some evidence has suggested that female participants draw on the language system more prominently than males [@hutchinson2013a], other research has suggested that this difference is negligible in the general population [@wallentinChapterGenderDifferences2020]. Presumably, the scarcity of statistical power that has ailed some studies in cognitive psychology and neuroscience [@vasishthHowEmbraceVariation2021; @marekReproducibleBrainwideAssociation2022] may have affected some studies in the present topic area as well [see @lynottReplicationExperiencingPhysical2014; @montero-melisNoEvidenceEmbodiment2022]. Therefore, in this study, we re-examine longstanding questions using larger-than-average data sets, and calculate the sample size required to reliably detect a range of effects.

## Language and vision

To retest the interplay theory, we revisit previous studies that contained larger-than-average samples of participants. Each of our three studies is organised around one *hub* study. In Study 1, the hub is @hutchison2013a, which implements a semantic priming paradigm. In Study 2, the hub is @pexman2017a, which implements a lexical decision paradigm. In Study 3, the hub is @balota2007a, which implements a lexical decision paradigm. Upon this basis, we added word-level variables from other studies to tap into language [@mandera2017a; @wingfieldUnderstandingRoleLinguistic2022] and vision [@lynott2020a; @petilli2021a]. Additionally, we included several covariates---or nuisance variables---to allow a rigorous analysis of the effects of interest [@sassenhagenCommonMisapplicationStatistical2016]. These covariates comprised participant-specific variables (e.g., attentional control), lexical variables (e.g., word frequency) and word concreteness.

# Levels of analysis

Experimental data in psycholinguistics often consists of various levels, such as participants, words and tasks. A simultaneous examination of these levels should enhance our understanding of the evidence collected [@ostarekStrongInferenceResearch2021]---for instance, by shedding light on the distribution of explanatory power within and across levels. This multilevel approach is complementary to another approach in the field that seeks to test the causal contribution of different sources of information to conceptual processing---e.g., language [@ponari2018a], perception [@stasenkoWhenConceptsLose2014] and action [@speedImpairedComprehensionSpeed2017].

The three levels examined in this study are described below.

## Individual level

Variation  The role of individual differences in domains such as language, perception, mental imagery and physical experience [e.g., @davies2017a; @dils2010a; @fettermanFeelingWarmBeing2018; @holt2006a; @mak2019a; @miceliDifferencesRelatedAging2022; @pexman2018a; @vukovic2015a; @yap2012a; @yap2017a].^[According to @lamiellStatisticalThinkingPsychology2019, 'individual differences' is a misnomer in that the methods used to analyse individual differences (e.g, regression) are not participant-specific. We think that this observation is mitigated in the context of mixed-effects models (as used in our current study), which incorporate by-participant random intercepts and random slopes.] Furthermore, in some topic areas in which individual differences have not featured as prominently, the data have revealed a non-trivial role of individual differences [@montero-melisConsistencyMotionEvent2021; @kosIndividualVariationLate2012].

### Vocabulary size

Vocabulary size refers to the number of words a person can recognise out of a discrete set. Higher-vocabulary participants are expected to respond faster overall [@pexman2018a]. Some previous studies have found that the effect of vocabulary size was moderated by variables related to general intelligence, such as processing speed [@ratcliff2010a; @yap2012a]. Owing to those findings, and the recommendations from other studies [@james2018a; @pexman2018a; @sassenhagenCommonMisapplicationStatistical2016], the present studies included covariates of vocabulary size, where available. These covariates were measures of intelligence that were not vocabulary-based, namely, attentional control in Study 1 and information uptake in Study 2 (such a variable was not available for Study 3).

## Word level

*** to be edited *** Semantic or lexical information in words [e.g., @dedeyneVisualAffectiveMultimodal2021; @lam2015a; @lund1995a; @lund1996a; @lynott2020a; @mandera2017a; @petilli2021a; @pexman2017a; @santosPropertyGenerationReflects2011; @wingfieldUnderstandingRoleLinguistic2022]

- Variables in this study: language-based and vision-based information;

- Covariates: lexical variables and word concreteness;

  - Lexical variables: The lexical covariates were selected in each study out of the same 5 variables, which have been identified as important in previous research (Wingfield & Connell, 2022; see [\underline{General statistical analysis}](#general-statistical-analyses) for details).

  - Word concreteness: Adjusting for word concreteness was deemed important due to the pervasive effect of this variable across lexical-semantic tasks [@brysbaert2014a; @connellStrengthPerceptualExperience2012; @pexman2018a]. Furthermore, it has been recently suggested that the effect of word concreteness does not stem from perceptual simulation, but instead from modality-independent properties of words [@bottiniConcretenessAdvantageLexical2021].

Studies have operationalised the *language system* using measures that capture the relationships among words without explicitly drawing on any sensory or affective modalities. Two main types of linguistic measures exist: those based on text corpora---dubbed *word co-occurrence* measures [@bullinariaExtractingSemanticRepresentations2007; @petilli2021a; @wingfieldUnderstandingRoleLinguistic2022]---and those based on associations collected from human participants---dubbed *word association* measures [@de2016a; @de2019a]. Word co-occurrence is more purely linguistic, as word association captures more of the sensory and affective meaning of words [@dedeyneVisualAffectiveMultimodal2021]. In Studies 1 and 2 below, word co-occurrence measures are used to represent the language system at the word level. In Study 3, however, such a measure could not be implemented. The measurement of priming across consecutive trials was not possible due to the high prevalence of nonword trials throughout the experiment.

The embodiment system has been represented by measures that explicitly draw on perceptual, motor or affective modalities [@fernandinoDecodingInformationStructure2022]. The perceptual modalities used in studies have often matched the 5 prototypical senses---vision, hearing, touch, taste, smell [@bernabeu2017a; @bernabeu2021a; @louwerseTasteWordsLinguistic2011]---and, less often, interoception [@connellInteroceptionForgottenModality2018]. Yet, out of these senses, vision has been most frequently used in research [e.g., @bottiniConcretenessAdvantageLexical2021; @dedeyneVisualAffectiveMultimodal2021; @pearsonHeterogeneityMentalRepresentation2015; @petilli2021a; @yeeColorlessGreenIdeas2012]. The prevalence of vision in research could be due to the prevalence of this sense in the brain [@reillyEnglishLexiconMirrors2020] and in language [@lynott2020a; @winterVisionDominatesPerceptual2018]. Furthermore, for the purpose of examining individual differences along with word-level variables, vision is easier to tap into than other senses. For instance, the perception of visual motion can be measured using a standard computer screen [@rajanandaRandomdotKinematogramWebbased2018]. Owing to these reasons, the three studies reported below implement the visual modality to represent the embodiment system at the word level.

## Task level

*** to be edited ***

- *Variable in this study:* stimulus-onset asynchrony (SOA), present in Study 1 only.^[The names of all variables were slightly adjusted for this text to facilitate their understanding---for instance, by replacing underscores with spaces (see code scripts at [https://osf.io/ueryq](https://osf.io/ueryq)). One specific case deserves further comment. We use the formula of the 'stimulus-onset asynchrony' (SOA) in this paper, instead of the 'interstimulus interval' (ISI), as the SOA has been more commonly used in previous papers [e.g., @hutchison2013a; @pecherDoesPizzaPrime1998; @petilli2021a; @yap2017a]. The difference between these formulas is that the ISI does not count the presentation of the prime word [example equivalences between ISI and SOA are available in @lam2015a, and in Figure 1A of @dilolloDecouplingStimulusDuration2004]. In the current study, the presentation of the prime word lasts 150 ms. Thus, the 200-ms SOA is equivalent to an ISI of 50 ms, and the 1,200-ms SOA corresponds to an ISI of 1,050 ms [@hutchison2013a]. The use of either formula in the analysis would not change our results, as we recoded the levels of the factor as -0.5 and +0.5, and then $z$-scored those, following the advice of @brauer2018a. In our analyses ([https://osf.io/ueryq](https://osf.io/ueryq)), we used the ISI formula as it was the one present in the data set of @hutchison2013a (retrieved from [https://www.montana.edu/attmemlab/documents/all%20ldt%20subs_all%20trials3.xlsx](https://www.montana.edu/attmemlab/documents/all%20ldt%20subs_all%20trials3.xlsx)]).

In addition, the different paradigms applied across the three studies can be compared, albeit cautiously as the studies differ in other influential ways, such as the number of participants and stimulus words. The three tasks examined in the present studies are likely to reflect different degrees of semantic depth, in the following order [for background, see @connell2013a; @lam2015a; @ostarekTaskdependentCausalRole2017; @versaceImpactEmbodiedSimulation2021; @wingfieldUnderstandingRoleLinguistic2022].

1. **Semantic decision** (Study 2) likely elicits the deepest semantic processing, as the instructions of this task ask for a concreteness judgement. In this task, participants are asked to classify words as abstract or concrete.

>>>>

  This manuscript is a draft. Correspondence can be addressed to Pablo Bernabeu on pcbernabeu@gmail.com. All the materials needed to inspect and reproduce the present research---comprising documentation, data, analysis and results---are available at [http://doi.org/10.17605/OSF.IO/UERYQ](http://doi.org/10.17605/OSF.IO/UERYQ).

abstract: |
  Research has suggested that conceptual processing depends on both language-based and vision-based information. We tested this interplay at three levels of the experimental structure: individuals, words and tasks. To this end, we drew on three existing, large data sets that implemented the paradigms of semantic priming, semantic decision and lexical decision. We extended these data sets with measures of language-based and vision-based information, and analysed how the latter variables interacted with participants' vocabulary size and gender, and also with presentation speed in the semantic priming study. We performed the analysis using mixed-effects models that included a comprehensive array of fixed effects---including covariates---and random effects. First, we found that language-based information was more important than vision-based information. Second, in the semantic priming study---whose task required distinguishing between words and nonwords---, both language-based and vision-based information were more influential when words were presented faster. Third, a 'task-relevance advantage' was identified in higher-vocabulary participants. Specifically, in lexical decision, higher-vocabulary participants were more sensitive to language-based information than lower-vocabulary participants. In contrast, in semantic decision, higher-vocabulary participants were more sensitive to word concreteness. Fourth, we demonstrated the influence of the analytical method on the results. These findings support the interplay between language and vision in conceptual processing, and demonstrate the influence of measurement instruments on the results. Last, we estimated the sample size required to reliably investigate various effects. We found that 300 participants were sufficient to examine the effect of language-based information contained in words, whereas more than 1,000 participants were necessary to examine the effect of vision-based information and the interactions of both former variables with vocabulary size, gender and presentation speed. In conclusion, this power analysis reveals the need to increase sample sizes when conducting research on perceptual simulation and individual differences.

@@

output:
  papaja::apa6_pdf:
    extra_dependencies: ['flafter']   # Place tables and figures below where they are first mentioned in the text.

# Latex preamble specifying layout
header-includes: 
# Background on the first 4 lines: https://stackoverflow.com/a/70658799/7050882
  - \renewcommand\author[1]{}
  - \renewcommand\affiliation[1]{}
  - \authorsnames[1, 2, 2]{Pablo Bernabeu, Dermot Lynott, Louise Connell}
  - \authorsaffiliations{{Department of Psychology, Lancaster University, UK}, {Department of Psychology, Maynooth University, Ireland}}
  - \renewcommand{\topfraction}{.75}        # Reduce page margins
  - \renewcommand{\bottomfraction}{.4}      # Reduce page margins
  - \renewcommand{\textfraction}{.15}       # Reduce page margins
  - \renewcommand{\floatpagefraction}{.6}   # Reduce page margins
  - \setlength{\@fptop}{0pt}                # Vertically-align tables and figures to the top
  - \setcounter{topnumber}{3}               # Reduce page margins
  - \setcounter{bottomnumber}{2}            # Reduce page margins
  - \setcounter{totalnumber}{4}             # Reduce page margins
  - \usepackage{placeins}                   # Allow use of `\FloatBarrier` to position tables and figures
  - \usepackage{amsmath}                    # Maths formatting
  - \usepackage{upgreek}                    # Disable the italicisation of Greek letters
  - \usepackage{booktabs}                   # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{longtable}                  # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{array}                      # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{multirow}                   # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{wrapfig}                    # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{float}                      # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{colortbl}                   # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{pdflscape}                  # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{tabu}                       # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{threeparttable}             # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{threeparttablex}            # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage[normalem]{ulem}             # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{makecell}                   # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{xcolor}                     # Dependency for R package 'kableExtra' (table formatting)
  - \renewcommand\appendix{}                # Remove redundant 'Appendix' label
---

@@

library(papaja)       # APA-formatted manuscript
library(knitr)        # Document rendering
options(kableExtra.latex.load_packages = FALSE,   # instead, LaTeX packages loaded in header/preamble above
        knitr.graphics.rel_path = FALSE)          # avoid warning (https://github.com/yihui/knitr/issues/2119#issuecomment-1092411877)
knitr::opts_knit$set(root.dir = '../')            # move working directory to the project root
library(kableExtra)   # Tables
library(dplyr)        # Data wrangling
library(formattable)  # Format numbers
library(kableExtra)   # Table formatting (e.g., `pack_rows()`)
library(stringr)      # Text processing
library(ggplot2)      # Plots
library(GGally)       # Correlation plots
library(sjPlot)       # Model plots
library(RColorBrewer) # Colours in plots
library(patchwork)    # Combination of plots
library(Cairo)        # Allows use of special characters such as dashes in plots
library(magick)       # Image rendering
library(tikzDevice)   # Image rendering

@@

# Subset of the semantic priming study that included vision-based similarity.
# Code for data set below in 'semanticpriming/data' folder
semanticpriming_with_visualsimilarity = 
  read.csv('semanticpriming/data/subset_with_visualsimilarity/semanticpriming_with_visualsimilarity.csv')

@@

# Code for models below in 'semanticpriming/bayesian_analysis' folder
semanticpriming_summary_informativepriors_exgaussian = 
  readRDS('semanticpriming/bayesian_analysis/results/semanticpriming_summary_informativepriors_exgaussian.rds')
semanticpriming_summary_weaklyinformativepriors_exgaussian = 
  readRDS('semanticpriming/bayesian_analysis/results/semanticpriming_summary_weaklyinformativepriors_exgaussian.rds')
semanticpriming_summary_diffusepriors_exgaussian = 
  readRDS('semanticpriming/bayesian_analysis/results/semanticpriming_summary_diffusepriors_exgaussian.rds')

@@

# Code for models below in 'semanticdecision/bayesian_analysis' folder
semanticdecision_summary_informativepriors_exgaussian = 
  readRDS('semanticdecision/bayesian_analysis/results/semanticdecision_summary_informativepriors_exgaussian.rds')
semanticdecision_summary_weaklyinformativepriors_exgaussian = 
  readRDS('semanticdecision/bayesian_analysis/results/semanticdecision_summary_weaklyinformativepriors_exgaussian.rds')
semanticdecision_summary_diffusepriors_exgaussian = 
  readRDS('semanticdecision/bayesian_analysis/results/semanticdecision_summary_diffusepriors_exgaussian.rds')

@@

# Code for models below in 'lexicaldecision/bayesian_analysis' folder
lexicaldecision_summary_informativepriors_exgaussian = 
  readRDS('lexicaldecision/bayesian_analysis/results/lexicaldecision_summary_informativepriors_exgaussian.rds')
lexicaldecision_summary_weaklyinformativepriors_exgaussian = 
  readRDS('lexicaldecision/bayesian_analysis/results/lexicaldecision_summary_weaklyinformativepriors_exgaussian.rds')
lexicaldecision_summary_diffusepriors_exgaussian = 
  readRDS('lexicaldecision/bayesian_analysis/results/lexicaldecision_summary_diffusepriors_exgaussian.rds')

@@

Over the last two decades, research in the cognitive sciences has suggested that conceptual processing depends on both language and embodiment systems. That is, understanding words involves---on the one hand---lexical and semantic associations of an amodal kind, and---on the other hand---modality-specific associations within perceptual, motor, affective and social domains [@barsalouLanguageSimulationConceptual2008; @connell2019a; @davisBuildingSemanticMemory2021; @khatin-zadehStrongVersionsEmbodied2021; @viglioccoTheorySemanticRepresentation2009]. Studies addressing these systems have found that the language system is overall more prevalent in word processing, producing larger effect sizes [@banksLinguisticDistributionalKnowledge2021; @kiela2014a; @louwerse2015a; @lam2015a; @pecherDoesPizzaPrime1998; @petilli2021a]. More intricately, the roles of language and embodiment are modulated by the characteristics of individuals, words and tasks. For instance, people's individual experience with language is associated with differential effects relating to phonological, lexical and semantic features of words [@jaredSkilledAdultReaders2017; @pexman2018a; @yap2012a; @yap2017a; @yapIndividualDifferencesJoint2009]. Similarly, physical expertise and perceptual biases are associated with differences in the mental simulation of meaning [@beilockSportsExperienceChanges2008; @calvo-merinoActionObservationAcquired2005; @vukovic2015a]. Furthermore, the embodiment system is especially suited for the processing of concrete concepts---e.g., *red*, *building* [@jonesDistrubutionalSemanticsStill2022; @kousta2011a; @ponari2018a; cf. @borghiAbstractConceptsExternal2022]. Embodied information also becomes more important in the following conditions: (I) later in the time courses of word recognition [@bernabeu2017a; @louwerseNeurologicalEvidenceLinguistic2012; cf. @petilli2021a] and property generation [@santosPropertyGenerationReflects2011; @simmonsFMRIEvidenceWord2008], (II) when participants produce slower responses [@louwerseTasteWordsLinguistic2011], and (III) in tasks that elicit deeper semantic processing---e.g., semantic decision, as opposed to lexical decision [@ostarekTaskdependentCausalRole2017; @petilli2021a]. Last, research in computational linguistics has provided further support for the complementarity of language and embodied information, by revealing increased predictive performance when models are provided with perceptual information on top of text-based information [@NIPS2013_7cce53cf; @roadsLearningUnsupervisedAlignment2020].

In spite of the amount of evidence demonstrating the interplay between language and embodiment, there are four reasons to continue testing the interplay theory. First, the coexistence of several systems in a scientific theory must be thoroughly justified due to the value of simplicity [@galleseBrainConceptsRole2005; @tillmanHowSharpOccam2015]. This scrutiny is particularly necessary because the language system has consistently produced larger effect sizes than the embodiment system [@banksLinguisticDistributionalKnowledge2021; @kiela2014a; @louwerse2015a; @lam2015a; @pecherDoesPizzaPrime1998; @petilli2021a]. Therefore, it should be ruled out that the language system can suffice in all contexts.

Second, it is important to examine both language and embodiment across various levels of the experimental structure---namely, individuals (i.e., due to individual differences such as vocabulary size), words (i.e., lexical and semantic variables) and tasks (i.e., experimental conditions affecting, for instance, processing speed). Some studies have approached this comprehensive structure but there is still room to widen the scope. One of the findings revealed by cross-level analyses is the influence of word processing tasks on the importance of modality-specific information. For instance, @connellSeeHearWhat2014 found that the vision-based information in words is important both for word identification (i.e., lexical decision) and for reading aloud (i.e., naming). In contrast, the auditory information in words is important for reading aloud but not so much for word identification. Another finding from cross-level research is a 'task-relevance advantage' for individuals that have a greater linguistic experience. Specifically, @pexman2018a found that higher-vocabulary individuals were more sensitive to task-relevant information, such as word concreteness in the semantic decision task. Furthermore, regarding embodiment, research has revealed that individuals who are briefly exposed to a certain sport develop neural activity that allows them to mentally simulate sport-specific actions during language processing [@beilockSportsExperienceChanges2008]. While these works have covered a large swathe of the present topic, one question remains unanswered: how does an individual's linguistic experience relate to their sensivity to both linguistic and embodied information in words?

Third, there is some inconclusive evidence. First, some findings have suggested that higher-vocabulary participants are more sensitive to language-based information---as reflected in greater semantic priming [@yap2017a]---, whereas other findings have suggested the opposite [@yapIndividualDifferencesJoint2009]. Second, some studies have suggested that the language system is activated before the embodiment system [@lam2015a; @louwerseTasteWordsLinguistic2011], whereas a recent study suggested that this pattern does not hold in the lexical decision task [@petilli2021a]. Third, some evidence has suggested that female participants draw on the language system more prominently than males [@burman2008a; @hutchinson2013a; @jung2019a; @ullman2008a], whereas other research has suggested that this difference is negligible in the general population [@wallentinChapterGenderDifferences2020].

Fourth, some of the previous studies could have been affected by the scarcity of statistical power that has been identified in cognitive psychology and neuroscience [@marekReproducibleBrainwideAssociation2022; @lynottReplicationExperiencingPhysical2014; @montero-melisNoEvidenceEmbodiment2022]. Problematically, low-powered studies present more errors in the estimation of effect sizes and $p$ values [@lokenMeasurementErrorReplication2017; @heymanReliabilityItemlevelSemantic2018; @vasishthStatisticalSignificanceFilter2018]. The current studies address these four key issues.

# The present studies {#present-studies}

We revisit three larger-than-average studies to investigate the interplay between language and embodiment in conceptual processing. We devote a study to each of the three original studies. Thus, Study 1 is centred on @hutchison2013a and uses the semantic priming paradigm. Study 2 is centred on @pexman2017a and uses the semantic decision paradigm. Study 3 is centred on @balota2007a and uses the lexical decision paradigm. Each of these central studies contained measures of participants' vocabulary size and gender. Furthermore, the core data sets were expanded by adding variables that captured the language-based information in words [@mandera2017a; @wingfieldUnderstandingRoleLinguistic2022] and the vision-based information in words [@lynott2020a; @petilli2021a]---the latter being used to represent the embodiment system. One of the key questions we investigated using this array of variables was whether individual differences in vocabulary and gender modulated participants' sensivity to the language-based and vision-based information in words. Alongside the effects of interest, several covariates were included in the models to allow a rigorous analysis [@sassenhagenCommonMisapplicationStatistical2016]. These covariates comprised measures of general cognition and lexical characteristics of the stimulus words. Last, in each study, we performed a statistical power analysis to help estimate the sample size needed to investigate a variety of effects in future studies.

Below, we delve into the language and the embodiment components of these studies.

## Language {#language}

Studies have operationalised the language system at the word level using measures that capture the relationships among words without explicitly drawing on any sensory or affective modalities. Two main types of linguistic measures exist: those based on text corpora---dubbed *word co-occurrence* measures [@bullinariaExtractingSemanticRepresentations2007; @petilli2021a; @wingfieldUnderstandingRoleLinguistic2022]---and those based on associations collected from human participants---dubbed *word association* measures [@de2016a; @de2019a]. Notwithstanding the interrelation between word co-occurrence and word association [@planchueloNatureWordAssociations2022], co-occurrence is more purely linguistic, whereas association indirectly captures more of the sensory and affective meaning of words [@dedeyneVisualAffectiveMultimodal2021].

### Operationalisation and hypotheses

In Study 1 (semantic priming) and Study 2 (semantic decision), co-occurrence measures were used to represent the language system at the word level. Specifically, in Study 1, this measure was called `language-based similarity`, and it was based on the degree of text-based co-occurrence between the prime word and the target word in each trial [@mandera2017a]. In Study 2, the measure was called `word co-occurrence`, and it was based on the degree of text-based co-occurrence between each stimulus word and the words 'abstract' and 'concrete' [@wingfieldUnderstandingRoleLinguistic2022]. In Study 3 (lexical decision), a co-occurrence measure could not be used, as the co-occurrence of words in consecutive trials could not be calculated due to the high frequency of nonword trials throughout the lexical decision task. Therefore, a single-word measure had to be used instead. Word frequency was used as it was the lexical variable, among five, that had the largest effect (see [\underline{Appendix A}](#appendix-A-lexical-covariates)).

At the individual level, language was represented by participants' vocabulary size in Studies 1 and 2, and by participants' vocabulary age in Study 3. Vocabulary *size* and *age* did not differ in any consequential way. They both captured the amount of vocabulary knowledge of each participant, by testing their knowledge of a small sample of pre-normed words, and thereby inferring their overall knowledge.

We hypothesised that word co-occurrence, word frequency and vocabulary size would all have facilitatory effects on participants' performance, with higher values leading to shorter RTs [@pexman2018a; @yapIndividualDifferencesJoint2009; @wingfieldUnderstandingRoleLinguistic2022].

## Embodiment represented by vision-based information

In previous studies, the embodiment system has been represented at the word level by perceptual, motor, affective or social variables [@fernandinoDecodingInformationStructure2022; @viglioccoTheorySemanticRepresentation2009; @wangSocialEmotionDimensional2021]. For instance, the perceptual modalities have often corresponded to the five Aristotelian senses---vision, hearing, touch, taste and smell [@bernabeu2017a; @bernabeu2021a; @louwerseTasteWordsLinguistic2011]---and, less often, to interoception [@connellInteroceptionForgottenModality2018]. Yet, out of all these domains, vision has been most frequently used in research [e.g., @bottiniConcretenessAdvantageLexical2021; @dedeyneVisualAffectiveMultimodal2021; @pearsonHeterogeneityMentalRepresentation2015; @petilli2021a; @yeeColorlessGreenIdeas2012]. The hegemony of vision is likely due to the central position of vision in the human brain [@reillyEnglishLexiconMirrors2020] as well as in several languages [@bernabeuDutchModalityExclusivity2018; @chenMandarinChineseModality2019; @miceliPerceptualInteroceptiveStrength2021; @morucciAugmentedModalityExclusivity2019; @roqueVisionVerbsDominate2015; @speedDutchSensoryModality2021; @speedGroundingLanguageNeglected2020; @lynott2020a; @vergallitoPerceptualModalityNorms2020; @winterVisionDominatesPerceptual2018; @zhongSensorimotorNormsChinese2022]. In the present study, we focussed on vision alone due to three reasons. First, we wanted to use a single variable to represent sensorimotor information, just as a single variable would be used to represent linguistic information. Using a single variable for each system facilitates the analysis of interactions with other variables. Second, vision is very prominent in cognition, as we just reviewed. Third, we had planned to use the present research to determine the sample size of a subsequent study that focusses on vision (indeed, the present study grew out of a statistical power analysis).

### Operationalisation and hypotheses
